An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-07 paper.tex
Update Introduction

This gives a unique opportunity to explore the development of the Internet
through its standards. Standards development is an inherently social and
political process \cite{RFC2026,RFC7282}, requiring cooperation and
consensus among a growing number of stakeholders. Due to the importance
of the Internet, it's critical to gain a coherent understanding of the
activities that take place within the IETF, to understand who develops
Internet standards and what factors may predict the success of an idea
or proposal for a new standard.

In this paper, we explore the operation of the IETF to understand who
develops Internet standards, how the participants interact, and what
factors may lead to success.
Section \ref{sec:background} introduces the IETF and the Internet Standards
development process, and reviews the data with which we work.
Section \ref{sec:trends-documents} considers trends in document production,
complexity, and correctness.
Section \ref{sec:trends-demographics} moves on to consider the people
who are involved in Internet standards development and how the demographics
of the community have shifted over time.
Section \ref{sec:org-dyn} studies the interaction between participants,
explores who has influence and impact in the IETF and how that influence is
evident in their activities, and the dynamics of communications within the
IETF.
Section \ref{sec:language} considers communications within the IETF from
the perspective of language to explore whether there are differences in
linguistic patterns and language use in more influential participants, and
to consider how the use of language reflects the consensus-driven nature of
the standards development process.
Section \ref{sec:success-factors} then considers factors that lead to
success, exploring what makes a document likely to be published as an RFC
and what makes a successful participant.
Finally, we review related work in Section \ref{sec:related} and conclude
in Section \ref{sec:conclusions}.

>>>>

The availability of this data gives a unique opportunity to explore the
development of the Internet. Standards development is an inherently social
and political process \cite{RFC2026,RFC7282}, requiring cooperation and
consensus among a growing number of stakeholders. Due to the importance
of the Internet, it's critical to gain a coherent understanding of the
activities that take place within the IETF, to understand who develops
Internet standards and what factors may predict the success of an idea
or proposal for a new standard.

In this paper, we explore the operation of the IETF to understand who
develops Internet standards, how the participants interact, and what
factors may lead to success:
\begin{itemize}
  \item Section \ref{sec:background} introduces the IETF and the Internet
    Standards development process, and reviews the available data.

  \item Section \ref{sec:trends-documents} considers trends in the
    production, complexity, and correctness of IETF standards documents
    over time.

  \item Section \ref{sec:trends-demographics} moves on to consider the
    people who are involved in Internet standards development and how the
    demographics of the community have shifted over time.

  \item Section \ref{sec:org-dyn} studies the interaction between
    participants, explores who has influence and impact in the IETF and how
    that influence is evident in their activities, and the dynamics of
    communications within the IETF.

  \item Section \ref{sec:language} considers communications within the IETF
    from the perspective of language to explore whether there are
    differences in linguistic patterns and language use in more influential
    participants, and to consider how the use of language reflects the
    consensus-driven nature of the standards development process.

  \item Section \ref{sec:success-factors} discusses what factors lead to
    success, exploring what makes a document likely to be published as an
    RFC and what makes a successful author.

\end{itemize}

Finally, we review related work in Section \ref{sec:related} and draw some
conclusions in Section \ref{sec:conclusions}. This paper is a summary and
collation of our prior published work \cite{mcquistin:2021:characterising,
khare:2022:web-we-weave, healey:2023:power, mcquistin:2023:errare,
khare:2023:tracing, karan:2023:leda, healey:2023:power-frontiers,
barnes:2024:temporal} that is submitted to stimulate discussion.
